---
title: "IST 5535 Lab 1"
author: "Mark Chafin"
date: "1/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Read Data

Use the read.csv() function to read the data into R. Make sure strings are read as factors. Call the loaded data college. Show the structure of the data.

```{r}
college <- read.csv("College.csv", stringsAsFactors = TRUE)
str(college)
```
```{r}
head(college)
```
The first column is just the name of each university. We can set college names as row names then remove the first column from the dataset.

```{r}
rownames(college) <- college[,1]
college$X <- NULL
head(college)

```
## 3.	Data Exploration
a.	Show summary statistics of the dataset.
```{r}
summary(college)
```

b.	Use pairs() method to create a scatterplot matrix of the first ten variables.
```{r}
pairs(college[,1:10])
```


c.	Draw a boxplot of Outstate versus Private. 
```{r}
plot(Outstate ~ Private, data = college)
```


d.	Create a new qualitative variable, called Elite, from the Top10perc variable. We divide universities into two groups based on whether or not the proportion of students coming from the top 10 % of their high school classes exceeds 50%.Then use the summary() function to see how many elite universities there are.

```{r}
college$elite <- ifelse(college$Top10perc > 50, "Yes", "No")
college$elite <- factor(college$elite)
summary(college)
```


e.	Draw a 2 by 2 plot which includes four histograms of four variables including Enroll, Outstate, Grad.Rate, and Books.
```{r}
par(mfrow=c(2,2))

hist(college$Enroll)
hist(college$Outstate)
hist(college$Grad.Rate)
hist(college$Books)
```


f.	Use ggplot2 package to visualize the relationship between graduation rate and room and board costs. What is the relationship? Does the relationship change when compare between elite and non-elite universities?

```{r}
library(ggplot2)

ggplot(data = college,aes(x=Room.Board, y=Grad.Rate)) +
  geom_point() +
  geom_smooth(method = 'lm', se=F) +
  labs(x="Room and Board Costs", y="Graduation Rate")


```

Does the relationship change when compare between elite and non-elite universities?

```{r}
ggplot(data = college,aes(x=Room.Board, y=Grad.Rate, col=elite)) +
  geom_point() +
  geom_smooth(method = 'lm', se=F) +
  labs(x="Room and Board Costs", y="Graduation Rate")
```




